---
title: "Class11"
author: "Max Strul"
format: pdf
editor: visual
toc: true
---

#Data

```{r}
url <- "https://bioboot.github.io/bggn213_W19/class-material/rs8067378_ENSG00000172057.6.txt"
data <- read.table(url)

```

#Q13: 
Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

```{r}
#factor(data[,2])
ag <- sum(data[,2]=="A/G")
gg <- sum(data[,2]=="G/G")
aa <- sum(data[,2]=="A/A")
ag
gg
aa
ag_exp <- data$exp[data[,2]=="A/G"]
gg_exp <- data$exp[data[,2]=="G/G"]
aa_exp <- data$exp[data[,2]=="A/A"]
median(ag_exp)
median(gg_exp)
median(aa_exp)
summary(data)

```


#Q14:
Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
library("ggplot2")

ggplot(data)+
  aes(geno,exp,col=geno)+
  geom_boxplot()+
  geom_point(position="jitter")
```
```{r}
t.test(data$exp[data$geno=="A/G"],data$exp[data$geno=="A/A"])

t.test(data$exp[data$geno=="A/G"],data$exp[data$geno=="G/G"])

t.test(data$exp[data$geno=="G/G"],data$exp[data$geno=="A/A"])
```
All groups are statistically significantly different from one another and it does appear that an SNP can alter expression levels of (ORMDL3)
