---
title: "Class19Miniproject"
format: pdf
editor: visual
---

# R Info

```{r}
library("datapasta")
library("dplyr")
```

# Data scraping

```{r}
cdc <- data.frame(
                                 Year = c(1922L,1923L,1924L,1925L,
                                          1926L,1927L,1928L,1929L,1930L,1931L,
                                          1932L,1933L,1934L,1935L,1936L,
                                          1937L,1938L,1939L,1940L,1941L,1942L,
                                          1943L,1944L,1945L,1946L,1947L,
                                          1948L,1949L,1950L,1951L,1952L,
                                          1953L,1954L,1955L,1956L,1957L,1958L,
                                          1959L,1960L,1961L,1962L,1963L,
                                          1964L,1965L,1966L,1967L,1968L,1969L,
                                          1970L,1971L,1972L,1973L,1974L,
                                          1975L,1976L,1977L,1978L,1979L,1980L,
                                          1981L,1982L,1983L,1984L,1985L,
                                          1986L,1987L,1988L,1989L,1990L,
                                          1991L,1992L,1993L,1994L,1995L,1996L,
                                          1997L,1998L,1999L,2000L,2001L,
                                          2002L,2003L,2004L,2005L,2006L,2007L,
                                          2008L,2009L,2010L,2011L,2012L,
                                          2013L,2014L,2015L,2016L,2017L,2018L,
                                          2019L),
         No..Reported.Pertussis.Cases = c(107473,164191,165418,152003,
                                          202210,181411,161799,197371,
                                          166914,172559,215343,179135,265269,
                                          180518,147237,214652,227319,103188,
                                          183866,222202,191383,191890,109873,
                                          133792,109860,156517,74715,69479,
                                          120718,68687,45030,37129,60886,
                                          62786,31732,28295,32148,40005,
                                          14809,11468,17749,17135,13005,6799,
                                          7717,9718,4810,3285,4249,3036,
                                          3287,1759,2402,1738,1010,2177,2063,
                                          1623,1730,1248,1895,2463,2276,
                                          3589,4195,2823,3450,4157,4570,
                                          2719,4083,6586,4617,5137,7796,6564,
                                          7405,7298,7867,7580,9771,11647,
                                          25827,25616,15632,10454,13278,
                                          16858,27550,18719,48277,28639,32971,
                                          20762,17972,18975,15609,18617)
       )
```


# ploting 

# Q1
```{r}
library("ggplot2")

casebyyear <- ggplot(cdc) + aes(Year,No..Reported.Pertussis.Cases)+
  geom_point()+
  geom_line()+
  labs(title="Pertussis Cases by Year 1922-2019",xlab="Year",ylab="Number of cases")
casebyyear
```
# Q2

```{r}
casebyyear+(geom_vline(xintercept=c(1946,1996)))
```
# Q3 What happened after introduction of the Ap vaccine?

We see an increase in number of cases. There was a celar decline after the introduction of the WP vaccine, and an increase after the switch to the aP vaccine

# Q3

```{r}
library("jsonlite")

subject <- read_json("https://www.cmi-pb.org/api/subject", simplifyVector = TRUE) 

head(subject, 3)


```
# Q4 How many aP and wP infancy vaccinated subjects are in the dataset?

```{r}
table(subject$infancy_vac)
```


# Q5. How many Male and Female subjects/patients are in the dataset?

```{r}
table(subject$biological_sex)
```

# Q6. What is the breakdown of race and biological sex (e.g. number of Asian females, White males etc…)?

```{r}
table(subject$ethnicity)
table(subject$race)

```


```{r}
library("lubridate")

```

# Q7 Using this approach determine (i) the average age of wP individuals, (ii) the average age of aP individuals; and (iii) are they significantly different?

```{r}
subject$age <- time_length(today()-ymd(subject$year_of_birth),'years')

subject$age_of_boost <- time_length(ymd(subject$date_of_boost)-ymd(subject$year_of_birth), 'years')
wp_ind <- subject %>% filter(infancy_vac=="wP")
ap_ind <- subject %>% filter(infancy_vac=="aP")

mean(wp_ind$age)
mean(ap_ind$age)
t.test(wp_ind$age,ap_ind$age)
```


# Q8. Determine the age of all individuals at time of boost?

```{r}
head(subject$age_of_boost)
```


# Q9. With the help of a faceted boxplot or histogram (see below), do you think these two groups are significantly different?

```{r}
ggplot(subject) +
  aes(time_length(age, "year"),
      fill=as.factor(infancy_vac)) +
  geom_histogram(show.legend=FALSE) +
  facet_wrap(vars(infancy_vac), nrow=2) 
```


# Dyplr function of *join

`inner_join()` joins everything that is the same row
`full_join()` this allows two datasets to join and keep everything 

We want to "join" the `subject` and `specimin` tables to have all the meta data we need tofr later analysis. We can use the dyplr function `*_join` function for this task. 

# Q10

```{r}
specimen <- read_json("https://www.cmi-pb.org/api/specimen", simplifyVector = TRUE) 


meta <- inner_join(specimen, subject)
titer <- read_json("https://www.cmi-pb.org/api/ab_titer", simplifyVector = TRUE) 
head(titer)
abdata <- inner_join(meta,titer)
```
# Q11. How many specimens (i.e. entries in abdata) do we have for each isotype?

```{r}
table(abdata$isotype)
```

# Q12. What do you notice about the number of visit 8 specimens compared to other visits?

```{r}
table(abdata$visit)
```
Its much smaller!

```{r}
ig1 <- abdata %>% filter(isotype == "IgG1", visit!=8)
head(ig1)
```
# Q13

```{r}
abdataboxplot <- ggplot(abdata)+
  aes(MFI, antigen) + geom_boxplot()
abdataboxplot
```

```{r}
abdataboxplot_norm <- ggplot(abdata)+
  aes(MFI_normalised, antigen) + geom_boxplot()


ggplot(ig1)+aes(MFI,antigen) +geom_boxplot()+facet_wrap(vars(visit),nrow=2)
```



```{r}
url <- "https://www.cmi-pb.org/api/v2/rnaseq?versioned_ensembl_gene_id=eq.ENSG00000211896.7"

rna <- read_json(url, simplifyVector = TRUE) 
```


```{r}
ssrna <- inner_join(rna, meta)
```



